### 2018-02-12 manuscript.tex
added new references, updated local copy of the manuscript, and changed figure 4 in the plots code.

\documentclass[9pt,twocolumn,twoside]{gsajnl}
% Use the documentclass option 'lineno' to view line numbers
\usepackage{caption}
\usepackage{subcaption}
\usepackage[nameinlink,capitalise]{cleveref}
\graphicspath{{figures/}}
\newcommand{\G}{\textbf{G }}

\articletype{inv} % article type
% {inv} Investigation 
% {gs} Genomic Selection
% {goi} Genetics of Immunity 
% {gos} Genetics of Sex 
% {mp} Multiparental Populations

\title{Linkage disequilibrium can drive strong and unstable variances and covariances in fitness-associated traits}

\author[$\ast$,1]{Kevin Gomez}
\author[$\dagger$]{Jason Bertram}
\author[$\dagger$]{Joanna Masel}

\affil[$\ast$]{Graduate Interdisciplinary Program in Applied Mathematics, University of Arizona, and}
% * <masel@email.arizona.edu> 2017-10-24T23:03:12.117Z:
%
% ^.
\affil[$\dagger$]{Department of Evolution and Ecology, University of Arizona}

\keywords{Keyword; Keyword2; Keyword3; ...}

\runningtitle{GENETICS Journal Template on Overleaf} % For use in the footer

%% For the footnote.
%% Give the last name of the first author if only one author;
% \runningauthor{FirstAuthorLastname}
%% last names of both authors if there are two authors;
% \runningauthor{FirstAuthorLastname and SecondAuthorLastname}
%% last name of the first author followed by et al, if more than two authors.
\runningauthor{Gomez \textit{et al.}}

\begin{abstract}
Abstract.
\end{abstract}

\setboolean{displaycopyright}{true}

\begin{document}

\maketitle
\thispagestyle{firststyle}
\marginmark
\firstpagefootnote
\correspondingauthoraffiliation{Please insert the affiliation correspondence address and email for the corresponding author. The corresponding author should be marked with a `1' in the author list, as shown in the example.}
\vspace{-11pt}%

\section{Introduction}

Natural selection acts on multiple traits simultaneously. The mean trait value in a population can change either because of direct selection on trait X, or because of selection on trait Y plus a genetic correlation between X and Y \citep{lande1979quantitative,lande1983measurement}. The role of genetic correlations in the response of quantitative traits to natural selection is described by Lande's equation ($\Delta \bar{z} = \G \hspace{.05in}\beta$), where $\Delta \bar{z}$ is the expected change in the population mean of a trait $\Delta \bar{z}$, $\beta$ is the selection gradient, and the components of the \G  matrix are the additive genetic covariances between traits. Lande's equation governs the short-term evolution of correlated quantitative traits under natural selection. Genetic variances and covariances can also evolve over appreciable time scales, and as a result, Lande's expression alone cannot tell us how correlated quantitative traits will change in the long-run.

When the \G matrix is stable, it is possible to infer past selection gradients and extrapolate future trait evolution (both direction and rate) from historical patterns. For example, studies of past macro-evolution in populations and species have used net selection gradients to draw connections between morphological evolution and reproductive isolation \citep{price1984genetic}, infer morphological and phylogenetic relations \citep{schluter1984morphological}, and determine the extent of local adaptation between populations \citep{dudley1996}. Similar work has looked the impact of genetic constraints on rates of adaptation \citep{teplitsky2014assessing,teplitsky2011quantitative}, their ability to bias phenotypic evolution over millions of years \cite{Schluter1996}.

The importance of \G's stability in evolutionary inferences such as these has motivated a large body of empirical work focused on measuring and comparing it between populations. Comparative studies over the decades have used different metrics and statistical procedures to check whether populations have similar \G matrices. Each of these metrics vary in their ability to detect conserved structure in genetic variances and covariances \cite{Steppan2002}. Matrix correlations and maximum liklihood tests, for example, often test for correlations between \G matrices or and their components, but do not give additional information about other important characteristics that can be shared between \G matrices. Methods based on the use of principle components, such as common principle component analysis (CPCA), in contrast, can compare features between \G matrices such as shape, size, and orientation. These methods have gained popularity for this reason, and because they are capable of revealing which evolutionary forces may have significantly influenced the past dynamics of the population's \G matrix \cite{Phillips1999,Arnold2008}.

A significant number of comparative studies examining either natural populations or laboratory experiments show that at least some features of the \G matrix are relatively stable. In over 70\% of comparisons between conspecific groups and species find that \G matrices have similar shape and orientation \cite{Arnold2008}. These include populations that have diverged over millions of years \cite{Arnold1999}, and across distinct environments \cite{puentes2016similarity,teplitsky2014assessing}, suggesting that aspects of genetic constraints can be maintained over significant time scales and across a diverse set of environmental factors. There are also comparative studies that find no related structure between \G matrices \cite{wilkinson1990resistance,roff1999does,Steppan1997,Waldmann2000}, and others that reveal large instabilities in \G over very short time scales  \cite{bjorklund2013rapid,Siren2017}. 
% \cite{roff1999does} \cite{Begin2003,Begin2004}

Quantitative genetic models of genetic variances and covariances provide some indication of how ecological and evolutionary forces contribute to stability in \G's dynamics \citep{Turelli1988,barton1989evolutionary,reeve2000predicting}. For example, changes in genotype $\times$ environment interactions directly affect genetic variances and covariances and lead to significant changes in \G over few generations. Other critical stability conditions involve the strength of selection and mutational covariances. Further analytical insights from these models require knowledge of unknown genetic details, such as the distribution of allelic effects \cite{burger2000mathematical}. Simulations have been used to bypass this problem and examine the dynamics of \G in more detail \cite{Jones2003,Jones2007}. In both theoretical analysis and numerical simulations, selection and mutations play an important role in ensuring the stability of the \G matrix, but the effects of the latter are largely assumed by most quantitative geneticists to be the result of pleiotropic effects \cite{lande1979quantitative,lande1980genetic,Arnold2008}(I have more citations, and will insert them).

The genetic correlations described by \G are due to some combination of pleiotropy and linkage disequilibrium \cite{Saltz2017}. Historically, pleiotropy was favored by the Edinburgh school, who argued that recombination would quickly eliminate linkage disequilibrium \citep[Chapter~20]{fox2006evolutionary}, particularly in the randomly mating animal populations that the Edinburgh school focused on. In contrast, quantitative geneticists in the alternative, Birmingham school largely worked on inbred lines of plants, where the effects of linkage disequilibrium were impossible to ignore. The Edinburgh view came to be widely accepted, and as a result, most models for the evolution of the \G  matrix assume that genetic correlations are pleiotropic and go on to examine how evolutionary processes influence them \citep{Turelli1988}. We know far less about the about the dynamics of genetic correlations arising from linkage disequilibrium.\par

To study linkage disequilibrium, we have to consider explicit alleles or genotypes, not just quantitative traits. Beneficial mutations are discrete not infinitesimal, and appear on distinct genetic backgrounds. Selective sweeps compete with one another, and so many beneficial mutations are lost in a process known as ``clonal interference''. This will slow down adaptation \citep{hill1966effect} unless recombination brings them together, reducing the negative linkage disequilibrium between competing beneficial mutations \cite{fisher1930genetical,muller1932some}. Classical models of clonal interference considered only two loci each with two alleles. More recently, traveling wave models have been developed to describe clonal interference among large numbers of loci, arising at an arbitrary mutation rate \citep{rouzine2003solitary,desai2007beneficial,park2010speed,good2012distribution,neher2010rate,fisher2013asexual,rouzine2010multi,rouzine2007highly,rouzine2008traveling}. These models have documented that clonal interference can have an enormous impact on adaptation rates. However, to date, most of these models have treated only the evolution of ``fitness'' rather than the evolution of individual traits and their correlations \cite{}. \par

% * <masel@email.arizona.edu> 2018-01-24T22:40:30.992Z:
% 
% > owever, to date, these models have treated only the evolution of ``fitness'' rather than the evolution of individual traits and their correlations. 
% You mentioned one paper that might be an exception - find it!
% 
% ^ <kgomez81@math.arizona.edu> 2018-02-10T00:01:15.252Z:
% 
% I am still looking for this reference.
%
% ^.
% * <masel@email.arizona.edu> 2018-01-24T22:37:04.241Z:
% 
% >  \citep{rouzine2003solitary,desai2007beneficial,park2010speed,good2012distribution}
% You can include sexual models here too, eg by Neher
% 
% ^ <kgomez81@math.arizona.edu> 2018-02-09T23:38:56.909Z:
% 
% I added the Neher paper, and other references. I'll conduct another search for other sources. 
% 
% ^.

Here, we focus on asexual populations, because they have the strongest clonal interference and resulting linkage disequilibrium. Moreover, biology today has come to recognize the numerical prevalence and ubiquitous significance of relatively asexual microbes \cite{mcfall2013animals}.\par

We begin with Desai and Fisher's \citep{desai2007beneficial} framework of fixed population size $N$, beneficial mutations at rate $U$, and per-beneficial-mutation selection coefficient $s$. This yields the rate of adaptation \citep[Equation 41]{desai2007beneficial}
\begin{equation}\label{eq:1}
v(U,N,s) = s^2\frac{2\log(N s)-\log(s/U)}{\log^2(s/U)}.
\end{equation}
Equation \ref{eq:1} holds when the parameters satisfy $NU \gtrsim 1/\ln(Ns)$ and $U/s \ll 1$ (the ``concurrent mutations regime''  where beneficial mutations appear rapidly relative to the time required for any one of them to fix). When combined with Fisher's fundamental theorem, Equation \eqref{eq:1} also gives the additive variance in fitness $\sigma^2$. When there is only one adaptive trait, $\sigma^2$ is that trait's additive variance. \par

With two adaptive traits, Equation \eqref{eq:1} still provides overall fitness variance $\sigma^2$, but it does not give its decomposition into the variances and covariance expressed in Lande's equation. Consider just two traits, each experiencing beneficial mutations at rate $U$ which each increase fitness by $s$. Equation \ref{eq:1} indicates that the overall adaptation rate is $v(2U,N,s)$, and so by symmetry, adaptation in each trait alone occurs at rate $0.5 \hspace{.02 in} v(2U,N,s)$. This is lower than the rate $v(U,N,s)$ when the other trait is not evolving. The reduction in trait-specific adaptation rate must be distributed between the two terms of Lande's equation $v_1 =\sigma_1^2 +\sigma_{1,2}$. More importantly, the contributions of each term will over time change if the \G matrix is not constant.\par

Here we analyze a two-dimensional traveling wave model of an asexual population, and find that clonal interference alone, in the absence of pleiotropy, leads on average to greatly elevated variance of fitness-associated traits and to negative covariance between them.  However, the \G matrix is highly unstable over the timescale of selective sweeps.

\section{Materials and Methods}
\label{sec:materials:methods}

We consider an asexual population of haploids with fixed total population size $N$ evolving in continuous time. To keep the model simple, we assume that beneficial mutations occur with the same rate of $U$ mutations per trait per birth giving a fitness benefit $s$ with no epistasis. Deleterious mutations are not considered. Since there is no epistasis, all individual with $i$ mutations in the first trait and $j$ in the second have the same Malthusian relative fitness $r_{i,j}= i s+j s$. We refer to each such set of individuals with the same $i$ and $j$ as a ``class'' and denote their abundances and frequencies by $n_{i,j}$ and $p_{i,j}$ respectively. The population's mean fitness is denoted $\bar{r}$.

We assume that $N$ is so large that selection is strong relative to drift ($1/N \ll s$). In this case, most classes are so large that their frequency grows or declines approximately deterministically due to selection. We refer to these classes as the ``bulk''  (Figure \ref{fig:1}, solid cells). However, adaptive mutant lineages arising along the high-fitness ``front'' of the population will start with a single individual, and initially behave stochastically (Figure \ref{fig:1}, patterned cells; arrows show mutations). Many of these lineages will go extinct before attaining appreciable frequencies, but some will make it to high enough abundances that they begin to grow deterministically. This transition from stochastic to deterministic behavior is called ``establishment'' and occurs if the lineage reaches an abundance of order $1/s$.  \par

The evolution of population follows an iterative process with two steps. First, the next class to establish is determined, and its depends on the growth of classes producing adaptive mutants in the front (\ref{fig:1}, arrows). Once a class has established, it begins to alter the growth dynamics of the other classes in the bulk. The second step involves producing the solutions for the dynamics of the bulk with the newly established class. Having done so, the process then repeats with determining the next class from the front to establish using the solutions obtained.

To simulate the timing of the next establishment, stochastic fluctuations induced by drift and mutations must be accounted for in the growth of classes at the front. These growth of classes in the front follow a branching processes fed by mutants from the adjacent (and exponentially growing) classes in the bulk.  Each of the mutant lineages that arise will either go extinct, or grow sufficiently large to experience deterministic growth from selection. The chances of the latter happening is given by the probability of establishment, denoted $\pi$. We use the value of $\pi$ given by the expression:
\begin{equation} \label{eq:2}
\pi = \frac{s_{i,j}}{ 1+s_{i,j}},
\end{equation}and derived from \cite[Equation 16]{desai2007beneficial}. Here $s_{i,j}=(r_{i,j}-\bar{r})$ is the selective advantage of an individual in the class with $i$ mutation in the first trait, and $j$ mutations in the second, with respect to an average individual in the population.  Although classes in the front can consist of many mutant lineages that contribute to their size during the stochastic growth phase, their establishment  is driven by the size of the first lineage destined to establish. In particular, the time required for this single lineage to transition into the deterministic growth phase is a good approximation for the actual time of establishment of its class.  The justification for this approximation is discussed in detail by \cite[p.1773]{desai2007beneficial}, and intuitively has to do with the fact that lineages from early mutation events often quickly die out. \par

The next establishment event and its corresponding class  is determined by sampling the each of the sojourn times of mutations in the front that are destined to establish and identifying the first to appear. Beneficial mutations within the bulk are ignored because these are lost to clonal interference. At the front, the number of beneficial mutations from a class $(i,j)$ in the bulk follows an inhomogeneous Poisson  process with a rate $\lambda_{i,j}(t)$. The rate will depend on the birth rate of individuals in the class $(i,j)$, their time-dependent abundances, and the beneficial mutation rate (see appendix).  The sojourn time $\tau$ for the appearance of the first mutation destined to establish can be transformed into the random variable $x=\int_0^\tau \pi_{fix} \lambda_{i,j}(t) dt$ that has a standard exponential distribution. The sampling of $\tau$ is be done by generating a value for  $x$ and numerically solving for $\tau$ in the integral equation. This is repeated for each class of the bulk adjacent to the stochastic front, and the next class to establish is picked to be the one with the smallest sojourn time. \par

A class that is chosen to establish is immediately incorporated into the bulk with an initial size that reflects stochastic effects on the growth of the single mutant lineage that drives its establishment. If a mutant lineage that is destined to establish appears at time $t=0$, it requires $T$ generations to grow to size $1/s_{i,j}$. The variable $T$ is random due to the effects of drift on the growth of the lineage, but its distribution is known and given by \cite[Equation 12]{desai2007beneficial}.  Once this lineage establishes $T$ generations later, the time dependent growth of its class is approximately $n_{i,j}(t) =s_{i,j}^{-1} \exp[{s_{i,j}(t-T)}]$. We consider instead an initial size  $\nu = s_{i,j}^{-1}\exp[-s_{i,j}T]$, and equivalently regard $n_{i,j}(t) = \nu s_{i,j}^{-1} \exp[{s_{i,j}t}] $ as describing the deterministic growth of the class beginning at time $t=0$. Consequently, each class that establishes can be incorporated into the bulk from the moment at which mutation driving establishment appears, but its initial size will be $\nu$. The random variable $\nu$ can be sampled by using its cumulative distribution function:
\begin{equation}\label{eq:3}
P(\nu \leq \nu_0) = 1- e^{-\pi_{fix} \nu_0},
\end{equation}
which is obtained from the probability density function of  $T$ in \cite[Equation 12]{desai2007beneficial}.

Following establishment, we update the set of classes in the bulk and calculate solutions to their time-dependent growth. Initially, classes with less than one individual are regarded as having gone extinct and removed from the bulk. Abundances of the remaining class and the newly established one change at a rate of  $\dot{n}_{i,j}(t) =(r_{i,j}-\bar{r})n_{i,j}(t)$ due to selection. Initial values are the size of each class at the time of establishment. Changes in the set of frequencies $p_{i,j}$ obey the same ordinary differential equation because $N$ is constant. Solutions of the coupled system of ODE's are then obtained using an ODE solver (NDSolve in Mathematica) for a period of one thousand generations, and are then used to determine the next establishment event. The process then continues for the duration of the simulation. Note that in our implementation the population size $N(t)=\sum n_{i,j}(t)$ can deviate slightly from the value $N$ as establishments take place and classes in the bulk become extinct. To account for this and maintain a total population size as close to $N$ as possible, we instead use the ODE 
\[ \dot{n}_{i,j} = (r_{i,j}-\bar{r})n_{i,j} - \bar{r}\left (\frac{N(t)}{N}-1 \right)n_{i,j} \] 
given by \cite[p.133]{Crow1970} in our simulations.\par

Simulations were initialized with monoclonal populations at carrying capacity consisting of individuals having one beneficial mutation in either trait and relative fitness set to zero. We varied either $N$, $s$ and $U$ while fixing the others. Each was varied across a range constructed from values reported by evolutionary experiments involving either E. coli or S. cerevisiae \citep{desai2007speed,Levy2015,Perfeito2007}.  We allowed the simulation to run for 5,000 generations, where one generation  This allowed the population to achieve beneficial mutation-selection balance before collecting data, and ensured that the transient effects from the initializing would not distort the statistics collected for the distribution. Following the burn-in period of the simulation, we recorded the existing classes in the bulk and their respective abundances. \par

The data was used to compute the marginal mean fitness and variances of each trait, as well as their covariance. From these quantities we also computed the instantaneous rate of adaptation (Equation \eqref{eq:5}) for each trait and overall.
The evolution of the population can be broken down into the evolution of the two traits, according to Lande's equation:
\begin{equation}\label{eq:5}
\left( \begin{array}{c}
\dot{\bar{r_1}} \\
\dot{\bar{r_2}} 
\end{array} \right)
=
\left( \begin{array}{cc}
\sigma_1^2 & \sigma_{1,2} \\
\sigma_{1,2} & \sigma_2^2 
\end{array}\right)
\left(\begin{array}{c}
1 \\
1 
\end{array}\right),
\end{equation}
where  $\bar{r_1}=\sum_{i,j} p_{i,j} is_1$, $\bar{r_2}=\sum_{i,j} p_{i,j} j s_2$, $\sigma_1^2 = \sum_{i,j} p_{i,j} (r_i-\bar{r_1})^2$, $\sigma_2^2 = \sum_{i,j} p_{i,j} (r_j-\bar{r_2})^2$, and  $\sigma_{1,2} =\sum_{i,j} p_{i,j} (r_i-\bar{r_1})(r_j-\bar{r_2})$. Here, it is the frequencies corresponding to the bulk that are relevant, since classes in the front are assumed to be too small to affect the means, variances and covariances of the distribution  ($1/s \ll N$).   Lande's equation above is expressed as the change in the population's mean trait values over time intervals of size $s$. \par

\section*{Results}
\label{sec:results}

The reduction in trait-specific adaptation $v_1$ from $v(U,N,s)$ down to $0.5 \hspace{.02 in} v(2U,N,s)$ (Figure \ref{fig:2}, extent to which solid lines are below 1), caused by clonal interference with the second trait, is driven by high levels of negative covariance (Figure \ref{fig:2}, closed circles). This is partly compensated for by a substantial increase in additive genetic variance (Figure \ref{fig:2}, open diamonds), because the negative covariance slows the removal of additive variance from the population.

In other words, clonal interference in a rapidly adapting population can explain the common observation of high genetic variance in a fitness-associated trait combined with strong negative covariance with other fitness-associated traits, even in the complete absence of pleiotropic trade-offs. This is striking, because this kind of evidence is often interpreted as evidence for trade-offs. \par

The reduction in $v_1$ due to the effects of clonal interference is insensitive to changes in the values of $s$ (Figure \ref{fig:2}a, solid lines),  $U$ (Figure \ref{fig:2}b), or $N$ (Figure \ref{fig:2}c). Generalizing to $k$ traits, each with equal $U$ and $s$, from Equation \eqref{eq:1}  we derive
\begin{equation}\label{eq:6}
 \tilde{v}_1 (U,s,N;k) 
= \frac{1}{k}\frac{\log^2[s/U]\log[N^2s(kU)]}{\log^2[s/(kU)]\log[N^2sU]},
\end{equation}
in units of multiples of the rate of adaptation in isolation, $v(U,N,s)$. Each additional trait increases clonal interference on the focal trait by a diminishing amount, with curvature apparent even with respect to the logarithm of the number of traits (Figure \ref{fig:3}). This suggests that much can be learned even from the simplest case of clonal interference between only two fitness-associated traits.\par

Under beneficial mutation-selection balance, the components of the \G matrix fluctuate far more than the instantaneous rate of adaptation, with transient spikes in additive genetic variance $\sigma_1^2$ mostly canceled out by transient spikes in negative covariance $\sigma_{1,2}$ (representative simulation shown in Figure \ref{fig:4}). As a result, fluctuations in variance and covariance can be substantially larger than those of total variance in fitness.\par

The two-dimensional traveling wave of Figure \ref{fig:1} can be projected onto a one-dimension traveling wave of fitness, where classes along a fitness isocline such as those along the red line in Figure \ref{fig:1}, are aggregated into fitness classes (Figure \ref{fig:4}). The shape of the one-dimensional fitness distribution is Gaussian and highly peaked near the population's mean fitness \citep{desai2007beneficial}. This means that the \G matrix will dominated by the distribution of genotypes in this fitness class (red line in Figure \ref{fig:1}, mode in (Figure \ref{fig:4}). As adaptation continues, a fitter class becomes the most abundant that will have a different distribution of genotypes and hence different trait variances and covariance. On average, the time required for this shift is the mean time between establishments at the nose of the of the traveling wave, denoted $\tau_q$ in \citet{desai2007beneficial}.

\begin{equation}\label{eq:8}
\tau_q = \frac{\log^2(s/U)}{s(2\log(Ns)-\log(s/U))}
\end{equation}\par

Genotypes frequencies within the peak of the traveling wave are set during the stochastic growth phase of the fitness class prior to establishing \citep{Desai2013}.  As previously discussed, beneficial mutations that give rise to mutants in the nose produce distinct but equally fit mutant lineages. Many of lineages become extinct, but those that manage to survive are subsequently and equally amplified by selection following establishment. Since the relative sizes of the surviving lineages with respect to one do not change once they begin growing deterministically, the frequencies of genotypes in a fitness class become ``frozen'' at the time of establishment. Beneficial mutations that occur after that time contribute very little to the size and composition of fitness class, since their corresponding lineages will be exponentially smaller than those which initially appeared. After a fitness class has established, it eventually goes on to become the dominant group of the fitness distribution, or rather, it forms the peak of the traveling wave. The time between its establishment and its succession as the dominant fitness class is called the sweep time (denoted $\tau_{sw}$). Consequently, the frequencies of genotypes in the nose determine the values of variance and covariances approximately $\tau_{sw}$ generations later. Figure \ref{fig:6} plots the correlation between trait covariances of the bulk and nose, as a function of a time offset. The signal peaks at approximately $\bar{\tau}_{fix}$ (mean time to fixation). It shows that covariances between traits are largely set by the behavior at the nose of the one dimensional traveling wave, which consists of genotypes along the lead fitness isocline through the stochastic front.  The values of variances and covariance will be correlated from one transition to the next, because mutational influx depends on the abundances of adjacent less-fit genotypes.

% (Dynamics of structures withing the \G matrix)
Principle components can be used to examine the structural features of our \G matrix. Although Fluctuations in the individual components $\sigma_1^2$, $\sigma_1^2$, $\sigma_{1,2}$ can be large relative to the rate of adaptation, but there sum has lower overall variance (Figure \ref{fig:5}, black line). The  
\[
E\left[
\left( \begin{array}{cc}
\sigma_1^2 & \sigma_{1,2} \\
\sigma_{1,2} & \sigma_2^2 
\end{array}\right)
\right] = 
\]

\begin{figure}[!ht]
\includegraphics[width=1\linewidth]{fig1a.pdf}
\caption{Representative two-dimensional genotype distribution. Individuals with equal numbers of beneficial mutations in each trait are combined into classes (boxes). Abundances of the bulk (solid boxes) behave deterministically, while abundance of the stochastic front (patterned boxes) behave stochastically. The red dot marks the average number of beneficial mutations in each trait, and equally fit classes are found along the red fitness isocline). In asexuals, beneficial mutations must occur on the fittest genetic backgrounds in order to contribute to the adaptive process. These mutations are represented by arrows from the bulk into the stochastic front. Classes in the stochastic front become part of the bulk once their abundances have become sufficiently large, pushing the stochastic front to include new classes. As classes below the population mean fitness isocline decline and those above increase exponentially, a two-dimensional traveling wave is produced. (Simulation parameters: $N=10^9$, $s=0.02$, and $U=10^{-5}$)}\label{fig:1}
\end{figure}

\begin{figure*}[!ht]
\centering
\includegraphics[width=1\linewidth]{fig2.pdf} 
\caption{Variance (open diamonds) of one trait and its covariance (closed circles) with the other, averaged across simulations, whose expected sum is equal to the analytically calculated adaptation rate (solid line). Variance of the focal trait is always higher than it would be in the absence of the second trait ($\sigma_1^2 >1$); negative covariance more than cancels this out ($|\sigma_{1,2}| > \sigma_1^2-1$) to reduce the adaptation rate.  Mean variance and covariance are insensitive to changes in the population size $N$ (c), but exhibit some dependence on the values of $s$ (a) and $U$ (b). For the parameter values not being varied on the x-axis, $s=0.02$, $U=10^{-5}$ and $N=10^9$. }\label{fig:2}
% * <masel@email.arizona.edu> 2018-01-22T19:36:50.877Z:
% 
% > averaged across simulations
% How many simulations were averaged?
% 
% ^.
\end{figure*}

\begin{figure}[!ht]
\centering
\includegraphics[width=1\linewidth]{fig3.pdf}
\caption{Rate of adaptation of a focal trait as a function of the number of total traits undergoing adaptation. Each additional trait reduces the rate of adaptation, but by marginally less as the trait number grows. The decrease decays to zero in the Equation \eqref{eq:6}, but in practice the expression is only valid when $kU/s \ll 1$. The effects of clonal interference on trait evolution are most significant in the case of two traits. The values of the parameters are taken as $s=0.02$, $U=10^{-5}$ and $N=10^9$.}
% * <masel@email.arizona.edu> 2017-10-28T22:59:26.669Z:
% 
% > Each plot includes the measured values produced by simulations with either varying $s$, $U$, or $N$. 
% Aren't var and covar from simulations and v from the equation?
% 
% ^.
\label{fig:3}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=1\linewidth]{fig4.pdf}
% * <masel@email.arizona.edu> 2017-10-28T23:28:47.376Z:
% 
% > fig3.pdf
% Need tick labels on the x-axis.
% 
% ^.
\caption{Trajectories of the variance of trait one (dashed line), covariance (dotted line), and the total variance in fitness of the population (solid line), plotted over 10,000 generations. The mean values of each are included as lines with matching line-styles. The \G matrix over this period of time is unstable. Its two components, trait one variance and covariance, fluctuate significantly relative to the variance in total fitness over identical time periods. The trajectories were sampled from one simulation with parameters $s=0.02$, $U=10^{-5}$ and $N=10^9$.}\label{fig:4}
% * <masel@email.arizona.edu> 2017-10-28T23:32:29.075Z:
% 
% > along with variance in total fitness
% I thought you were plotting sigma_1^2+sigma_12 which is the instantaneous adaptation rate of trait 1, but here you say that you are plotting something else. Please confirm.
% 
% ^.
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=0.49\linewidth]{fig5a.pdf}
\includegraphics[width=0.49\linewidth]{fig5b.pdf}
\caption{The dynamics of the \G matrix components are governed by time scale $\tau_q$ of the one-dimensional traveling wave of fitness classes. Variances and covariance are dominated by the set of genotypes that make up the peak of the one-dimensional traveling wave, whose shape is approximately Gaussian. At time $t$ the peak is composed of classes shown in blue, whose frequencies determine the components of the \G  matrix. At time $t+\tau_q$ the mean shifts (relatively abruptly, due to the exponential growth of classes fitter than the mean), and the set of genotypes shown in green become the dominant group, whose distribution dominates the new trait variances and covariance.}\label{fig:5}
% * <masel@email.arizona.edu> 2017-10-28T23:57:21.626Z:
% 
% > whose shape is approximately Gaussian
% The text hasn't really explained or provided a citation for this point.
% 
% ^.
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=1\linewidth]{fig6a.pdf}
\caption{Correlation between covariance of stochastic front and bulk.  The correlation between the two peaks when their signals are offset by the the mean time to fixation $\bar{\tau}_{fix}$ (marked by the blue line), indicating that the dynamics of the stochastic front explains 69\% of the covariance lated detected in the bulk $\bar{\tau}_{fix}$ generation after.}\label{fig:6}
% * <masel@email.arizona.edu> 2017-10-28T23:57:21.626Z:
% 
% > whose shape is approximately Gaussian
% The text hasn't really explained or provided a citation for this point.
% 
% ^.
\end{figure}

\section*{Discussion}

% * <masel@email.arizona.edu> 2018-01-17T17:50:06.655Z:
% 
% > Discussion}
% Include speculation about the effects of adding deleterious mutations.
% 
% ^ <masel@email.arizona.edu> 2018-01-24T22:30:26.120Z:
% 
% Compare to all other theoretical sources of G-matrix instability eg. mutation drift migration etc. and argue it is strong relative to others.
%
% ^.
\label{sec:discussion}
% (Genetic Constraints on the Rate of Adaptation)
% \cite{phillips1989visualizing} discusses a unifying framework for micro- and macro-evolution based on a unification of Lande's quantitative genetics theory \cite{lande1979quantitative} and Simpson's adaptive landscapes \cite{simpson1944tempo}. His discussion centers on the role of the |G matrix, which encodes genetic constraints, in defining the local behavior of evolutionary trajectories on the fitness landscape. Aspects of this framework has been tested emprically by many. \cite{Schluter1996} analysis of empirical data on threespine stickleback populations showed that phenotypic responses over the past four million were biased towards \G's leading eigenvector ($g_{max}$); similar patterns were found in data on mice, sparrows, and finches. Studies like these have prompted subsequent investigations of the impact that genetic constraints might have on the rate of adaptation. \cite{teplitsky2014assessing} for example, estimated population-specific \G matrices and multivariate selection coefficients form long-term data for seven bird species (10 populations). He found that genetic correlations reduced the rate of adaptation by 28\% on average, and multivariate evolvability tended to be systematically lower in the direction of selection compared to others (see also Teplitsky et al. 2011). \cite{kopp2014rapid} provides a review of this subject in relation to the potential for rapid adaptation to environmental changes given genetic constraints, phenotypic plasticity, and evolvability, based on different models. Their discussion of genetic constraints in quantitative genetic models review most of the finding expressed in past papers (see Schluter 1996, Phillips et al 1989, Hansen and Houle 2008). It is still unclear whether genetic constraints do in general limit the rate of adaptation, since other related studies have not been so conclusive on the issue \citep{Agrawal2009,hine2014evolutionary}. In addition, different metrics have been used measure the impact of genetic constraints on rates of adaptation using \G \citep{HANSEN2008,Kirkpatrick2009,Agrawal2009}, and there subtle differences in what they indicate about constraints (see Kopp et al. 2014). Some of the issues may in fact be related to the use of \G itself \citep{chebib2017affects}, which may not truly measure genetic constraints at all. The role of genetic constraints must be addressed. \cite{matuszewski2015catch} showed in their work that standing genetic variation has greater potential for facilitating adaptation to environmental change than de novo beneficial mutations. In their approach, which combined those of adaptive walks models (discrete substitutions from new mutations) and quantitative genetic models (adaptation from standing genetic variation), and they found that populations that adapt from standing genetic variation are capable of significantly more phenotype change over time, and thus, have a higher potential to adapt when faced with environmental change that is fast, rather than slow.

% ((General theory for changes in G))
Quantitative genetic models for the evolution of \G indicate that selection, drift, mutation and recombination each induce changes in genetic variances and covariances over time. Lande's Gaussian framework in particular is often used to specify a decomposition of changes in the \G matrix into three parts represented in the expression:
\[\Delta G = G(\gamma - \beta \cdot \beta^T)G + 2 M + 2\sum_{i,j=1}^n r_{i,j}(C_{i,j}-C_{i,j}'),\]
which is given in \cite[Chapter 20, Box 3]{fox2006evolutionary}. The first term describes the action of selection, which includes directional selection eroding genetic variation ($\beta$ terms), as well as stabilizing and correlational selection which reorienting the G matrix ($\gamma$ term). The second term involves the mutational covariance matrix \textbf{M}, and gives the sum of mutational covariances generated by the effects of pleiotropic mutations at each locus. The third term gives the rate of degradation in covariance caused by linkage disequilibrium, which depends on the recombination rate. More terms may be needed depending on whether the distribution of allelic effects are normal (Barton and Turelli 1987). More detailed explanations of each terms are given in  \cite[Chp34,35,36]{lynch2018evolution}, but the last term is discussed under assumptions that are valid for infinitesimal model.

% (Overview of Genetic Drift and its relevance in our model)
With respect to the importance of mutations, pleiotropy is largely emphasized \cite{reeve2000predicting,Steppan2002,Arnold2008}. Changes in genetic correlaitons due to linkage disequilibria enter into the dynamics of \G through the mutation matrix \textbf{M}. 
\cite{Phillips2001}
\cite{lande1979quantitative}

% (Fluctuations in G due to Mutations)
\cite{Arnold2008} has a section on stability of the G matrix when mutational effects evolve (Jones papers cited). 
\cite{Jones2003} The nature of mutations at pleiotropic loci can dramatically influence stability of G, i.e. when a mutation at a single locus simultaneously changes the value of the two traits (due to pleiotropy) and these effects are correlated, mutation can generate extreme stability of G. 
\cite{Jones2007} examine a two trait system and find that mutational correlation does evolve in response to selection on the bivariate phenotype (mutation is third quant-trait).\par

% Quantitative genetic models for the dynamics of \G attribute changes in size are thought to be driven by drift \cite{lande1979quantitative,Roff2000,Arnold2008}. In our model, changes in the size of the \G matrix are induced by stochastic fluctuations in the front resulting from the effects of drift.  
% (Fluctations in G: Nose Mutations vs Mutation Matrix)
% \cite{Jones2007} 
% \cite{Arnold2008}
% \cite{HANSEN2008}

% (Genetic correlations as constraints to trait evolution)
% \cite{hine2014evolutionary}

% (Comparison with Barton and Otto's Negative Linkage Disequilibria Results, and other related quantitative genetics)
% \cite{Barton2005} 
% \cite{Bulmer1971} Selection induces changes in variance contributions from linkage disequilbrium are shaped by selection, since the infinitesimal model, frequencies 
% \cite{}
% In our model, covariance between traits is arises entirely from linkage disequilibrium. We can assume that specific alleles at either locus correspond to the number of beneficial mutations that have accumulated in the corresponding trait.  It follows that the $p_{i,j}$ are the gametic frequencies for the genome consisting of allele $i$ paired with allele $j$, and $D_{i,j} = p_{i,j}-p_{i\cdot} p_{\cdot j}$ is their coefficient of linkage disequilibrium ( $p_{i\cdot}$ and $p_{\cdot j}$ are marginal distributions). By applying \textbf{equation 2}, we can derive a mathematical relationship between covariance and linkage disequilibrium (need to cite chapter 5, Walsh and Lynch).
% % * <masel@email.arizona.edu> 2017-07-29T17:34:47.794Z:
% % 
% % > specific alleles at either locus correspond to the number of beneficial mutations that have accumulated in the corresponding trait
% % Multiple loci can contribute to the same trait. An allele cannot be a "number of beneficial mutations", because an allele must pertain to one locus. Stick to the word "genotype" not "allele", eg "genotype i,j has i beneficial mutations in trait 1 and j beneficial mutations in trait 2 compared to the ancestral population". This means that you cannot calculate linkage disequilibrium as you do: linkage disequilibrium is generally defined as being between two particular beneficial mutations. What you calculated was simply a covariance. Because of this confusion, I got a bit lost here and so skipped Equation 6.
% % 
% % ^.
% \begin{equation}
% \sigma_{1,2}=\sum_{i,j}D_{i,j}[(i-\bar{i})s_1][(j-\bar{j})s_2]. 
% \end{equation} 
% As the population adaptsAcknowledgments, mutations and selection  increase and decrease the amounts LD which ultimately determine how traits interact in their evolution.\par

% We examined the resulting dynamics of trait evolution using numerical simulations of our model using Mathematica. We focused on modeling the symmetric case, in which the two traits had identical mutation rate and selection coefficients. These, along with the population size, were chosen to match values reported in evolutionary experiments involving relevant micro-organisms (see Table 1).

% The behavior of the stochastic front is an evolving boundary in two-dimensional trait space. \cite{pearce2017rapid} derive features of the steady state two-dimensional front in the asymptotic limit $\log(N s) \rightarrow \infty$, but these do not characterize its stochasticity. As we will later show, the dynamics variances and covariance depend significantly on these fluctuations. Thus, we proceed by simulation. [Polish]

\section*{Acknowledgments}
\label{sec:acknowledgments}
We thank Patrick Phillips and Bruce Walsh for helpful discussions. Funding was provided by the National Science Foundation (DEB-1348262) and the National Institutes of Health (T32 GM084905).
\bibliography{biblio}

\section*{Appendix}
The derivation of the expression given in \ref{eq:4} from Lande's equation is as follows. The derivative measuring the change in the mean of a trait $\dot{\bar{r_1}} \approx \Delta \bar{r_1}/\Delta t$ for a sufficiently small interval of time. Lande's original expression yields,
\[
\left[
\begin{array}{c}
\Delta \bar{r_1} \\
\Delta \bar{r_2} 
\end{array}
\right]
=
\left[
\begin{array}{cc}
\sigma_1^2 & \sigma_{1,2} \\
\sigma_{1,2} & \sigma_2^2 
\end{array}
\right]
\left[
\begin{array}{c}
\beta_1 \\
\beta_2 
\end{array}
\right]
\]
The selection gradient on the right hand side is the change in fitness with respect to change in the trait value, indicating that $\beta_1=\beta_2=s$. Dividing both sides by $\Delta t$ allows us to equate the left side to the column vector $(\dot{\bar{r_1}},\dot{\bar{r_2}})^T$, while the scaled selection gradient becomes $(s/\Delta t,s/\Delta t)$. Setting $\Delta t = s$ provides equation 5.

\section{Relevant That Needs Discussion}
in asexuals; this fact has, for example, been the basis of explanations for the evolutionary advantages of sex and recombination \citep{Barton2005,Otto2009}
\end{document}Acknowledgments